Dynamical Properties of Molecular Cloud Complexes in Galaxies at the Epoch of Reionization

### 2019-06-25 WRITEUP/ms.tex
updated Fig 6

\begin{figure*}[htbp]
 \centering
  \includegraphics[width=\textwidth, trim=25 25 35 30, clip]{\figpath/SS_bigfonts/{combined}.pdf}
\caption{
Examples of MCCs (white contours) identified by the clump finder in
\flower during its accreting phase (\Fig{SFH}a; left two columns), where \flower display a highly disturbed gas morphology, and 
its quiescent phase (right two columns), where \flower shows a rotating disk configuration.
Panels in the second and fourth columns show the molecular gas distribution projected along different viewing angles. 
Colorbar on the right shows the mean H$_2$ number density, weighted by gas mass. 
Different rows show the results obtained by applying different H$_2$ number density cuts $(n_{\rm cut})$ as annotated by the label in the bottom left corners. 
\label{fig:MCC}
}
\end{figure*}

@@

The most massive ($M_{\rm gas}>$10$^7$\,\Msun) cloud corresponds to the molecular disk of the galaxy (see e.g., \Fig{MCC}), 
which is identified as a single component at low $n_{\rm cut}$. This main disk component occupy the top right corner of \Fig{stellarRatio16}, with a \AP{do we really need both $f_{\rm gas}$ and $M_\star/M_{\rm gas}$?} 
stellar-to-gas mass ratio of $\simeq$\,50--100. 
The velocity dispersion of such structure is dominated by bulk motions, as shown in \Fig{vv}, where we compare 
the velocity dispersions of MCCs resulting from bulk ($\sigma_{\rm bulk}$) versus non-thermal turbulent motions ($\sigma_{\rm NT}$) 
in the accreting and quiescent phases of \flower.
The accreting phase of \flower displays a disturbed morphology, whereas the 
quiescent phase displays a disk-like morphology (\Fig{MCC}).
Excluding the structure corresponding to the molecular disk, the 
velocity dispersions of most MCCs in the accreting phase of \flower are dominated by turbulent motions.
On the other hand, $\sigma_{\rm bulk}$ is at least comparable to $\sigma_{\rm NT}$ for almost half the MCCs
in the quiescent disk-like phase.

@@

\begin{figure}
\centering
\includegraphics[trim=0 0 0 0, clip, width=0.5\textwidth]{\figpath/ss16-sigma-kms-bulk_sigma-kms-NT} \\
\includegraphics[trim=0 0 0 0, clip, width=0.5\textwidth]{\figpath/ss28-sigma-kms-bulk_sigma-kms-NT}
\caption{Gas velocity dispersions resulting from bulk versus turbulent motions for MCCs identified across all $n_{\rm cut}$ (same color-coding as \Fig{stellarRatio16}) in the accreting (top) and quiescent phase (bottom) of \flower.
%
The accreting phase of \flower displays a more disturbed morphology compared to the quiescent phase, which displays a disk-like morphology (\Fig{MCC}).
%
Velocity dispersion from turbulent motions dominates over bulk motions for most MCCs in the accreting phase. On the other hand, $\sigma_{\rm bulk}$ is at least comparable to $\sigma_{\rm NT}$ for almost half the MCCs in the quiescent disk-like phase.
\label{fig:vv}}
\vspace{0.5em}
\end{figure}

@@

% where the virial parameter is used as a surrogate for comparing the potential to kinetic energies of molecule gas structures. 
We perform virial analysis, as motivated by \obs, to assess the stability of MCCs. On average, the virial parameter of MCCs in the starburst phase of \flower is lower than in the accreting phase, as expected for \SF. Similarly, the virial parameter of MCCs identified at the highest density thresholds are lower than those identified at lower density thresholds. 
% Such MCCs are thus susceptible to collapse. 
% The high virial parameter found for some MCCs are found in regions of $Q_{\rm eff}>1$.
% On the other hand, 
Such MCCs identified in (the denser) regions with a lower Toomre $Q_{\rm eff}$ parameter are unstable against collapse.
% Toomre:
Close resemblance of $Q_{\rm eff}$ and $Q_{\rm star}$ maps indicates that contribution from the stellar component plays an important role in governing the stability of the MCCs against axisymmetric perturbations, especially in the central part of \flower. Similarly, stabilizing effect due to the thickness of its disk is also non-negligible. This illustrates the importance of accounting for stellar contribution and disk thickness when examining the stability of molecular gas structures, especially in relatively evolved and enriched systems at high redshift that are preferentially being observed now.
%
Star formation is expected to take place within its star-forming {\it clumps} and {\it cores} on smaller scales, where energy quickly dissipates. This is consistent with the notion that collapsing structures result from gravitational instability occurring within globally stable structures, which are supported by turbulence and rotation on large scale.
%%
%Low-$Q$ sets the regions that fragments 
% and within those regions where alphavir is low enough is where structures to further collapse to form stars.
%
%The MCCs of \flower are more massive and turbulent compared to the Milky Way, with 
% higher gas mass surface densities and velocity dispersions. 
This also implies that \obs on spatially resolution better than $\simeq$40\,pc are needed to examine the truly star-forming structures (cores), and thus, \SF in the first galaxies.
%
%% end
%Determining the multi-phase ISM properties of early galaxies
%is a critical piece to understanding the evolution and
%assembly history of galaxies, since they set the pace
%for chemical reactions and excitation rates for the coolants in the ISM (and subsequent star formation).
%While observations leveraging the combination of spatial-spectral imaging of
%multi-band continuum and spectral line emission are crucial,
%% we show in this work that prototypical galaxies at the EoR with properties similar to \flower
%% are expected to have molecular gas structures on sub-kpc scales. 
% Higher resolution imaging of molecular gas 
Such resolution is within reach with ALMA, for which 0.01" resolution images are available for the high freq bands. The Next Generation VLA (ngVLA) in the 2030s will enable deeper imaging of the cold gas down to $\sim$\,200\,pc via \aco line emission. % 0.005" highest freq; ~0.05 for 16 GHz; CO10 for z~6
On the other hand, cosmological zoom-in simulations, such as \ncode{Serra}, while inherently limited in galaxy statistics and are subject to the sub-grid models adopted, serve as a useful tool for examining and making predictions on the morphology and dynamics of the molecular ISM of the first galaxies.
%These predictions can in turn be
%tested with high resolution imaging of the gas content in the first galaxies with telescopes such as ALMA and the ngVLA, thereby
%enabling us to test our findings and refine future simulations to shed light on the physical processes behind
%\SF since the cosmic dark ages.

>>>>

\begin{figure*}[htbp]
 \centering
  \includegraphics[width=\textwidth, trim=25 25 35 30, clip]{\figpath/SS_bigfonts/{combined}.pdf}
\caption{
Examples of MCCs (white contours) identified by the clump finder in
\flower during its accreting phase (\Fig{SFH}a; left two columns), where \flower display a highly disturbed gas morphology, and 
its late quiescent phase (right two columns), where \flower shows a rotating disk configuration.
Panels in the second and fourth columns show the molecular gas distribution projected along different viewing angles. 
Colorbar on the right shows the mean H$_2$ number density, weighted by gas mass. 
Different rows show the results obtained by applying different H$_2$ number density cuts $(n_{\rm cut})$ as annotated by the label in the bottom left corners. 
\label{fig:MCC}
}
\end{figure*}

@@

The most massive ($M_{\rm gas}>$10$^7$\,\Msun) cloud corresponds to the molecular disk of the galaxy (see e.g., \Fig{MCC}), 
which is identified as a single component at low $n_{\rm cut}$. This main disk component occupy the top right corner of \Fig{stellarRatio16}, with a 
low gas mass fraction of $\simeq$\,0.
The velocity dispersion of such structure is dominated by bulk motions, as shown in \Fig{vv}, where we compare 
the velocity dispersions of MCCs resulting from bulk ($\sigma_{\rm bulk}$) versus non-thermal turbulent motions ($\sigma_{\rm NT}$) 
in the accreting and quiescent phases of \flower.
The accreting phase of \flower displays a disturbed morphology, whereas the 
quiescent phase displays a disk-like morphology (\Fig{MCC}).
Excluding the structure corresponding to the molecular disk, the 
velocity dispersions of most MCCs in the accreting phase of \flower are dominated by turbulent motions.
On the other hand, $\sigma_{\rm bulk}$ is at least comparable to $\sigma_{\rm NT}$ for almost half the MCCs
in the quiescent disk-like phase.

@@

\begin{figure}
\centering
\includegraphics[trim=0 0 0 0, clip, width=0.5\textwidth]{\figpath/ss16-28-veldisp.pdf}
\caption{Gas velocity dispersions resulting from bulk versus turbulent motions for molecular structures 
identified across all $n_{\rm cut}$, as shown using the same color-coding as \Fig{stellarRatio16}.
%
Top and bottom panels show structures identified in the accreting phase and the quiescent phase, respectively.
The accreting phase of \flower displays a more disturbed morphology compared to the quiescent phase, which displays a disk-like morphology (see \Fig{MCC}).
%
The dashed black lines show a one-to-one mapping between the axes to facilitate comparison. 
%
Velocity dispersion from turbulent motions dominates over bulk motions for MCCs in the accreting phase. On the other hand, 
$\sigma_{\rm bulk}$ is comparable to $\sigma_{\rm NT}$ for some MCCs in the quiescent disk-like phase.
\label{fig:vv}}
\vspace{0.5em}
\end{figure}

@@

% where the virial parameter is used as a surrogate for comparing the potential to kinetic energies of molecule gas structures. 
We perform virial analysis, as motivated by \obs, to assess the stability of MCCs. On average, the virial parameter of MCCs in the starburst phase of \flower is lower than in the accreting phase, as expected for \SF. Similarly, the virial parameter of MCCs identified at the highest density thresholds are lower than those identified at lower density thresholds. 
% Such MCCs are thus susceptible to collapse. 
% The high virial parameter found for some MCCs are found in regions of $Q_{\rm eff}>1$.
% On the other hand, 
Such MCCs identified in (the denser) regions with a lower Toomre $Q_{\rm eff}$ parameter are unstable against collapse.
% Toomre:
Close resemblance of $Q_{\rm eff}$ and $Q_{\rm star}$ maps indicates that contribution from the stellar component plays an important role in governing the stability of the MCCs against axisymmetric perturbations, especially in the central part of \flower. Similarly, stabilizing effect due to the thickness of its disk is also non-negligible. This illustrates the importance of accounting for stellar contribution and disk thickness when examining the stability of molecular gas structures, especially in relatively evolved and enriched systems at high redshift that are preferentially being observed now.
%
Star formation is expected to take place within its star-forming {\it clumps} and {\it cores} on smaller scales, where energy quickly dissipates. This is consistent with the notion that collapsing structures result from gravitational instability occurring within globally stable structures, which are supported by turbulence and rotation on large scale.
%%
%Low-$Q$ sets the regions that fragments 
% and within those regions where alphavir is low enough is where structures to further collapse to form stars.
%
%The MCCs of \flower are more massive and turbulent compared to the Milky Way, with 
% higher gas mass surface densities and velocity dispersions. 
This also implies that \obs on spatially resolution better than $\simeq$40\,pc are needed to examine the truly star-forming structures (cores), and thus, \SF in the first galaxies.
%
%% end
%Determining the multi-phase ISM properties of early galaxies
%is a critical piece to understanding the evolution and
%assembly history of galaxies, since they set the pace
%for chemical reactions and excitation rates for the coolants in the ISM (and subsequent star formation).
%While observations leveraging the combination of spatial-spectral imaging of
%multi-band continuum and spectral line emission are crucial,
%% we show in this work that prototypical galaxies at the EoR with properties similar to \flower
%% are expected to have molecular gas structures on sub-kpc scales. 
% Higher resolution imaging of molecular gas 
Such resolution is within reach with ALMA, for which 0\farcs01 resolution images are available for the high frequency bands. The Next Generation VLA (ngVLA) in the 2030s will enable deeper imaging of the cold gas down to $\sim$\,200\,pc via \aco line emission. % 0.005" highest freq; ~0.05 for 16 GHz; CO10 for z~6
On the other hand, cosmological zoom-in simulations, such as \ncode{Serra}, while inherently limited in galaxy statistics and are subject to the sub-grid models adopted, serve as a useful tool for examining and making predictions on the morphology and dynamics of the molecular ISM of the first galaxies.
%These predictions can in turn be
%tested with high resolution imaging of the gas content in the first galaxies with telescopes such as ALMA and the ngVLA, thereby
%enabling us to test our findings and refine future simulations to shed light on the physical processes behind
%\SF since the cosmic dark ages.
